### 2015-11-28 wiley-big-data-devel/table-1.tex

% Government-hosted databases Datasets table
\begin{table}[h!]
\caption{Illustrative big datasets hosted by governmental entities}
\tiny
\centering
\begin{tabular}{| p{2.1in} | p{2.1in} | p{2.1in} | } 
\hline
Dataset & Sample & URL \\
\hline\hline
Current Population Survey (CPS) & & \url{http://www.bls.gov/cps/demographics.htm} \\
\hline
Danish National Birth Cohort & n=100,000 women recruited 1999-2002, assessed during pregnancy and when children were 6 mos, 18 mos, 7 yrs, and 11 yrs & \url{http://www.ssi.dk/English/RandD/Research%20areas/Epidemiology/DNBC/} \\
\hline
Early Childhood Longitudinal Study (ECLS) & Birth cohort (ECLS-B): n=14,000 children born in 2001 from birth through kindergarten; ECLS-K cohort: kindergarten through 8th grade; ECLS-K:2011: kindergarten through 5th grade. & \url{http://nces.ed.gov/ecls/} \\
\hline
Japan Environment and Children`s Study (JECS) & n=100,000 parent-child pairs & \url{http://www.env.go.jp/en/chemi/hs/jecs/} \\
\hline
National Health and Nutrition Examination Survey (NHANES) & nationally-representative sample of n=5,000  & \url{http://www.cdc.gov/nchs/nhanes.htm} \\ 
\hline
National Longitudinal Survey of Youth (NLSY) & n=10,000 youth born 1957-1964 surveyed beginning in 1979. & \url{http://www.bls.gov/nls/} \\
\hline
National Youth Fitness Survey (NYFS) & n=1,650 3-15 year-olds surveyed in 2012 & \url{http://www.cdc.gov/nchs/nnyfs.htm} \\
\hline
NICHD Study of Early Child Care and Youth Development (SECCYD) & n=1,364 infants born in 1991 & \url{https://www.nichd.nih.gov/research/supported/Pages/seccyd.aspx} \\
\hline
NIH MRI Study of Normal Brain Development & n=554 4-18 year-olds & \url{http://pediatricmri.nih.gov/nihpd/info} \\
\hline
NIMH National Database for Autism Research (NDAR) & n>85,000 individuals & \url{http://ndar.nih.gov} \\
\hline
Norwegian Mother and Child Cohort Study (MoBa) & n>90,000 pregnant women and n>70,000 men, recruited 1999-2008 & \url{http://www.fhi.no/eway/} \\
\hline
WHO Multicentre Growth Reference Study (MGRS) & n=8,500 0-24 mo-olds and 18-71 mo-olds & \url{http://www.who.int/childgrowth/en/} \\
\hline
\end{tabular}
\label{table:1}
\end{table}