### 2015-11-28 wiley-big-data-devel/table-2.tex

% Datasets hosted by non-governmental entities
\begin{table}[h!]
\caption{Illustrative big developmental datasets hosted by non-governmental entities.}
\tiny
\centering
\begin{tabular}{| p{2.1in} | p{2.1in} | p{2.1in} | } 
\hline
Dataset & Sample & URL \\
\hline\hline
National Longitudinal Study of Adolescent Health (AddHealth) & Wave I: n=90,118 adolescents in-school questionnaires, n=20,745 in-home interviews; Wave II: n=14,738 adolescents in-home interviews; Wave III
n=15,197 young adults in-home interviews and biomarker collection; Wave IV:
15,701 adult in-Home interviews and biomarker collection. & \url{http://www.cpc.unc.edu/projects/addhealth}\\
\hline
Child Language Data Exchange System (CHILDES) & n>3,000 6 mos - 8 yrs & \url{http://childes.psy.cmu.edu/} \\
\hline
Colorado Adoption Project (CAP) & n=450 families (>2,400 individuals) & http://ibg.colorado.edu/cap/ \\
\hline
Databrary & >3,400 hours of video from n>2,900 participants, 2 mos - 50+ yrs; experimental displays & \url{http://databrary.org} \\
\hline
Developing Human Connectome Project & 20 to 44 weeks post-conception & \url{http://www.developingconnectome.org} \\
\hline
Family Life Project (FLP) & n=800 sampled at 2, 6, 15, 24, and 36 mos & \url{http://flp.fpg.unc.edu} \\
\hline
Genome of the Netherlands Project (GoNL) & n=250 familes & \url{http://www.nlgenome.nl} \\
\hline
Human Speechome Project & 10 hrs video/day from 1 child from 0-3 yrs & \url{http://www.media.mit.edu/cogmac/projects/hsp.html} \\
\hline
Maternal Lifestyle Study (MLS) & n=1,388 infants (and mothers/caregivers) assessed at 4, 8, 10, 12, 18, 24, 30, and 36 mos & \url{https://neonatal.rti.org/about/mls_background.cfm} \\
\hline
Measures of Effective Teaching Project (MET) & n>3,000 teachers recorded beginning in 2009 & \url{http://www.metproject.org} \\
\hline
MetroBaby & n>1,000 racially and ethnically diverse children video-recorded at home or in the lab at 14, 24, 36, 52, 64, and 76 months & (\url{http://doi.org/10.17910/B7CC74} \\
\hline
Panel Study of Income Dynamics (PSID) & n=5,000 families (>18,000 individuals) & \url{https://psidonline.isr.umich.edu} \\
\hline
Pediatric Imaging, Neurocognition, and Genetics (PING) & n=1,000 3-20 year-olds & \url{http://pingstudy.ucsd.edu} \\
\hline
Progress in International Reading Literacy (PIRLS) & & \url{http://timssandpirls.bc.edu} \\
\hline
Psychiatric Genomics Consortium (PGC) & n>900,000 & \url{http://www.med.unc.edu/pgc} \\
\hline
Trends in International Math and Science Study (TIMSS) & n>20,000 per sample & \url{http://timss.bc.edu} \\
\hline
Trends in International Math and Science Study Video (TIMSSVideo) & teachers recorded in n=7 countries & \url{http://www.timssvideo.com} \\
\hline
Twins Early Development Study (TEDS) & survey data from n=15,000 twins born in 1994-1996; lab-based behavioral task and DNA samples from a smaller subset & \url{http://www.teds.ac.uk} \\
\hline
Twin and Offspring Study in Sweden (TOSS) & target of n=900 parents and families (~3,000 individuals) & \url{http://ki.se/en/meb/twin-offspring-study-in-sweden-toss} \\
\hline
\end{tabular}
\label{table:2}
\end{table}