### 2015-11-28 wiley-big-data-devel/table-3.tex

\begin{table}[h!]
\caption{Illustrative measure-based datasets.}
\tiny
\centering
\begin{tabular}{| p{3.25in} | p{3.25in} | }
\hline
Measure & Comments \\
\hline\hline
Bayley Scales of Infant and Toddler Development & \cite{bayley2006bayley}. Data underlying published norms not available, but some study-specific data is available: \url{http://www.epicure.ac.uk} and \url{http://www.icpsr.umich.edu/icpsrweb/ICPSR/studies/4091}. \\
\hline
\hline
Computer-based cognitive tutor data & LearnLab (\url{http://learnlab.org/technologies/datashop/index.php}) hosts DataShop repository/data analysis tool. \\
\hline
Infant Behavior Questionnaire (IBQ) & \cite{rothbart1981measurement}. Information at \url{http://www.bowdoin.edu/~sputnam/rothbart-temperament-questionnaires/instrument-descriptions/infant-behavior-questionnaire.html}. IBQ data on n=1,388 participants archived at ICSPR as part of the Maternal Lifestyle Study (MLS; \url{https://neonatal.rti.org/about/mls_background.cfm}) \\
\hline
Laboratory Temperament Assessment Battery (Lab-TAB) & \cite{goldsmith1993laboratory}. Information available at \url{http://www.uta.edu/faculty/jgagne/labtab} \\
\hline
MacArthur Communicative Inventory (CDI) & n>40,000 samples archived at WordBank (\url{http://wordbank.stanford.edu}) \\
\hline
Spatial Learning & NSF-funded Spatial Intelligence and Learning Center (SILC; \url{http://spatiallearning.org}) stores information about research tests and intruments \\ 
\hline
Teller Acuity Cards & Publications citing norms for 0-4 year-olds have been published \cite{teller1986assessment, mayer_monocular_1995, salomao_large_1995} \\
\hline 
Wechsler Intelligence Scale for Children-III (WISC-III) & Designed for children ages 6-16. Data underlying norms not available. Published by Pearson (\url{http://www.pearsonclinical.com/psychology/products/100000771/wechsler-intelligence-scale-for-childrensupsupfifth-edition--wisc-v.html})\\
\hline
Wechsler Preschool and Primary Scale of Intelligence-R (WPPSI-R) & Designed for children aged 4-6 1/2 years. Data underling norms not available. \\
\hline
\end{tabular}
\label{table:3}
\end{table}