### 2016-04-30 nyu-data-science-reproducibility-16/video-reproducibility.Rmd
KEA comments

---
title: "Video Doesn't Lie"
author: "Karen E. Adolph, David Millman, and Rick O. Gilmore"
output: 
  ioslides_presentation:
    widescreen: TRUE
    css: css/databrary.css
    logo: img/databrary-leaf-large.png
bibliography: bib/databrary.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Reproducible workflows with [Databrary](http://databrary.org)

## Acknowledgments

- [NSF BCS-1238599](http://www.nsf.gov/awardsearch/showAward?AWD_ID=1238599&HistoricalAwards=false)
- [NICHD U01-HD-076595](https://projectreporter.nih.gov/project_info_description.cfm?aid=8531595&icde=15908155&ddparam=&ddvalue=&ddsub=&cr=1&csb=default&cs=ASC)
- [Society for Research in Child Development (SRCD)](http://srcd.org)
- NYU Libraries, ITS, IRB, and OSP

## Is psychological science reproducible?

<div class="centered">
<a href="http://cos.io">
<img src="https://cos.io/static/img/icons/cos_wide.png">
</div>
</a>

- Nosek et al. [[@collaboration_estimating_2015]](http://doi.org/10.1126/science.aac4716)
- vs. Gilbert et al. [[@gilbert_comment_2016]](http://doi.org/10.1126/science.aad7243)
- Sharing [data](https://osf.io/ezcuj/wiki/home/?_ga=1.260625677.1251838540.1458403228) focuses attention on substantive issues

----

<iframe src="http://databrary.org" width=640px height=400px>
</iframe>

## [Databrary.org](http://databrary.org)

- Digital data library specialized for research video
- Video/audio + participant/context metadata
- Restricted access for research/educational use
- Policy framework for sharing identifiable data

## Video uniquely captures behavior

<div class="centered">
<video width="640" height="480" controls>
  <source src="https://nyu.databrary.org/slot/9840/-/asset/11192/download?inline=true" type="video/mp4">
Your browser does not support the video tag.
</video>
[[@c7fbeef0-d8a9-4b99-9e17-f598507486b7]](http://doi.org/10.17910/B7H019)
</div>

## Video uniquely captures behavior

<div class="centered">

<video width="640" height="480" controls>
  <source src="https://nyu.databrary.org/slot/9840/-/asset/11193/download?inline=true" type="video/mp4">
Your browser does not support the video tag.
</video>
[[@c7fbeef0-d8a9-4b99-9e17-f598507486b7]](http://doi.org/10.17910/B7H019)
</div>

https://nyu.databrary.org/slot/9840/-/asset/11193/download?inline=true

## [[@c7fbeef0-d8a9-4b99-9e17-f598507486b7]](http://doi.org/10.17910/B7H019)

<div class=centered>
<a href="https://nyu.databrary.org/volume/33">
<img src="img/frank-databrary.jpg" height=450px>
</a>
</div>

## Unequivocally & unambiguously demonstrates phenomena

<div class="centered">
<a href="http://doi.org/10.1126/science.1093567">
<img src="img/deloache.science.jpg" height=350px>
</a>

[[@deloache_scale_2004]](http://doi.org/10.1126/science.1093567)
</div>

----

<div class="centered">
<video width="640" height="480" controls>
  <source src="https://nyu.databrary.org/slot/9850/-/asset/11550/download?inline=true" type="video/mp4">
Your browser does not support the video tag.
</video>
[[@b03c5c6c-31b8-445e-9e4f-8917b735bb53]](http://doi.org/10.17910/B7H019)
</div>

----

<iframe src="http://doi.org/10.17910/B7H019" width=640px height=400px>
</iframe>

## Video depicts unreported methodological details

Fall from video, bridge 20 going into bridge 14, then fall. Subject 1. Cari brown sleeveless shirt. John assist.
<https://nyu.databrary.org/volume/5>

## [Datavyu](http://datavyu.org) makes video coding reproducible

- Raw research video must be coded
- [Datavyu.org](http://datavyu.org) is a free, [open source](https://github.com/databrary/datavyu) coding tool
- Add codes, annotations time-locked to video segments
- Ruby API for [scripting](https://github.com/databrary/Datavyu-Example-Scripts)

## Reproducible video coding

Static image of fall from video.

## Securing permission to share while limiting access

<div class="centered">
<video width="640" height="480" controls>
  <source src="https://databrary.org/video/example-video-1.mp4" type="video/mp4">
Your browser does not support the video tag.
</video>
</div>

## Databrary enhances reproducibility

- "Active" curation of data **as it is collected**
- Organize, share, standardized participant metadata
- Sharing based on access level & participant permission

## [[@ddff6ac6-cd1b-4365-89aa-70ea654b4ed9]](http://dx.doi.org/10.17910/B7MW2K)

<iframe src="https://nyu.databrary.org/volume/5" width=750px height=500px>
</iframe>

## Databrary enhances reproducibility

> - Share raw research videos
> - Share video excerpts for teaching and learning
> - Share materials, coding spreadsheets, displays
> - Share links to papers, code repositories
> - Shared datasets get DOIs
> - Long-term preservation via partnership with NYU Libraries

## Take homes

- Video uniquely captures complexity & richness of behavior
- Databrary securely stores, shares video
- Databrary and Datavyu enhance reproducibility of video-based research
- Databrary makes psychological science more transparent & reproducible

## References {.smaller}