### 2016-05-02 nyu-data-science-reproducibility-16/be-bold.Rmd
add be bold

---
title: "To boldly go where no science..."
author: "Karen E. Adolph, David Millman, and Rick O. Gilmore"
output: 
  ioslides_presentation:
    widescreen: TRUE
    css: css/databrary.css
    logo: img/databrary-leaf-large.png
bibliography: bib/databrary.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Acknowledgments

- [NSF BCS-1238599](http://www.nsf.gov/awardsearch/showAward?AWD_ID=1238599&HistoricalAwards=false)
- [NICHD U01-HD-076595](https://projectreporter.nih.gov/project_info_description.cfm?aid=8531595&icde=15908155&ddparam=&ddvalue=&ddsub=&cr=1&csb=default&cs=ASC)
- [Society for Research in Child Development (SRCD)](http://srcd.org)
- NYU Libraries, NYU IT, IRB, and OSP

## TL;DR

- Video is essential for reproducible behavioral science
    + Even where video is *not* the primary raw data
- [Databrary](http://databrary.org) specializes in storing and sharing research video
- Video is identifiable, but it *can* be shared
- *All* behavioral scientists should collect and share videos of their research

## Is behavioral science reproducible?

- Nosek et al. [[@collaboration_estimating_2015]](http://doi.org/10.1126/science.aac4716)
- vs. Gilbert et al. [[@gilbert_comment_2016]](http://doi.org/10.1126/science.aad7243)

## Why we're unsure

- Behavior rich, complex
- Numeric, text-based measures reduce that complexity
- **Video captures and preserves it**

## Why we're unsure

- Replications can fail due to methdological differences
- Methods sections don't always report essential details
- **Video captures and preserves it**

## **Essentials** for a reproducible behavioral science

- Video record all tasks, measures, and behaviors
- Share the recordings
- Share all questionnaires, tasks, displays
- Share statistical, computational, data workflows
- Prepare to share from the beginning
- Seek permission to share data

## [Databrary.org](http://databrary.org)

- Digital data library specialized for research video
- Video/audio + participant/context metadata
- Restricted access for research/educational use
- Policy framework for sharing identifiable data
- Developmental science focus, but not exclusive

----

<iframe src="http://databrary.org" width=640px height=400px>
</iframe>

## How Databrary overcomes barriers to sharing video

- Policies for sharing identifiable video data
- Tools for coding video
- Tools for "active curation"" during data collection

## Policies

- Restrict access to authorized researchers (& affiliates)
    + Institutional [access agreement](https://databrary.org/access/policies/agreement.html)
- Seek permission to share data from participants
    + Release [template](https://databrary.org/access/policies/release-template.html); [script](https://databrary.org/access/guide/investigators/release/asking/script.html); sample [video](https://databrary.org/video/example-video-1.mp4)

## Standardized release levels

<div class="centered">
<a href="https://databrary.org/access/guide/investigators/release/release-levels.html">
<img src="img/release-levels.jpg">
</a>
</div>

## Tools for coding video

- Raw research video must be coded
- [Datavyu.org](http://datavyu.org) = free, [open source](https://github.com/databrary/datavyu) coding tool
- Add codes, annotations time-locked to video segments
- Ruby API for [scripting](https://github.com/databrary/Datavyu-Example-Scripts) reproducible workflows

----

<div class="centered">
<a href="http://datavyu.org/user-guide/index.html">
<img src="http://datavyu.org/user-guide/_images/annotated-datavyu-desktop.png" width=1000px>
</a>
</div>

## Curating data **as it is collected**

- Organize, share, standardized participant metadata
- Sharing based on 
    + user access level
    + participant permission

## [[@ddff6ac6-cd1b-4365-89aa-70ea654b4ed9]](http://dx.doi.org/10.17910/B7MW2K)

<iframe src="https://nyu.databrary.org/volume/5" width=750px height=500px>
</iframe>

----

<div class="centered">
<img src="img/vol-5-spreadsheet-0.jpg">
</div>

----

<div class="centered">
<img src="img/vol-5-spreadsheet-1.jpg">
</div>

----

<div class="centered">
<img src="img/vol-5-spreadsheet-2.jpg">
</div>

----

<div class="centered">
<img src="img/vol-5-spreadsheet-3.jpg">
</div>

----

<div class="centered">
<img src="img/vol-5-spreadsheet-4.jpg">
</div>

----

<div class="centered">
<img src="img/databrary-vol-5-materials.jpg">

[[@ddff6ac6-cd1b-4365-89aa-70ea654b4ed9]](http://dx.doi.org/10.17910/B7MW2K)
</div>

## Take homes

- Video uniquely captures complexity & richness of behavior
- Video captures vital methodological details in behavioral research
- Identifiable video (and other data) can be shared
- Video data sharing essential for a reproducible behavioral science

## **Essentials** for a reproducible behavioral science

- Video record all tasks, measures, and behaviors
- Share the recordings
- Share all questionnaires, tasks, displays
- Share statistical, computational, data workflows
- Prepare to share from the beginning
- Seek permission to share data

## Join us!

<iframe src="https://databrary.org/about/jobs.html" width=640px height=400px>
</iframe>

## References {.smaller}